---
title: "Self Organizing Maps (SOM) for spatial segregation"
date: 23/09/2020
output: 
  html_document:
      code_folding: show
      collapsed: false
      toc: no
      toc_float: 
         collapsed: false

header-includes:
   - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
   - \usepackage[utf8]{inputenc}
   - \usepackage{booktabs}
---

The document consist of reproducing de SOM methodology for cluster 6 zones in Conception Metropolitan Area, Chile.
```{r,warning=FALSE,message=FALSE,error=FALSE}
library(SOMbrero)
library(dplyr)
library(ggplot2)
library(markdown)
library(summarytools)
options(scipen = 999)
set.seed(67)
```
 Connecting with the data base
```{r}
ANALISIS<-read.csv('data/DB.csv')
```
 The number of neurons is chosen by
```{r}
ncel=round(sqrt(5*sqrt(dim(ANALISIS)[1])),digits=0)
```
 Neural network training
```{r,results='hide'}
training<-trainSOM(x.data=select(ANALISIS,PEJH12,TII,DBI,H,PHZ),
                   dimension=c(ncel,ncel),
                   radius.type='letremy',
                   verbose=TRUE,
                   maxit=5000)
```
 Evaluating quality model
```{r,warning=FALSE,message=TRUE}
summary(training)
plot(training,what ='obs',show.names = FALSE)
```
 Dendrograma analysis
```{r}
nclas=6
scEST_S7<-superClass(training,k=nclas)
plot(scEST_S7,lwd=1.5)
plot(scEST_S7,type='grid',plot.legend=TRUE,show.names = FALSE)
```
 Heat Maps
```{r,warning=FALSE}
plot(training, what="obs", type="color", variable=1, print.title=FALSE, main='Porcentaje Escolaridad del Jefe de hogar menor a 12') + geom_raster()
plot(training, what="obs", type="color", variable=2, print.title=FALSE, main='Tasa Inmigración Intrametropolitana') + geom_raster()
plot(training, what="obs", type="color", variable=3, print.title=FALSE, main='Distancias al Bicentro más cercano') + geom_raster()
plot(training, what="obs", type="color", variable=4, print.title=FALSE, main='Diversidad de Usos de Suelo') + geom_raster()
plot(training, what="obs",type="color", variable=5, print.title =FALSE, main='Porcentaje Hacinamiento Zonal') + geom_raster()
```
 Then what you need to know are the descriptive statistics by cluster. For that, you need to make arrangements before starting.
```{r}
ID   <-names(training$clustering)
NEUR <-as.integer(training$clustering)
CLUSTER<-rep(0,length(NEUR))
for(i in 1:length(NEUR)){ CLUSTER[i]<-scEST_S7$cluster[NEUR[i]] }
MERGE<-merge(data.frame(ID,NEUR,CLUSTER), select(ANALISIS, ID,PEJH12,TII,H,PHZ,DBI), by.x="ID", by.y="ID")
```
 With the latter, we have the neuron and the supercluster of each observation, so from now on you can carry out the process
```{r,echo=FALSE,message=TRUE}
summary(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==1))
summary(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==2))
summary(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==3))
summary(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==4))
summary(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==5))
summary(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==6))
```
Comparative boxplots
```{r}
boxplot(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==1)$PEJH12,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==2)$PEJH12,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==3)$PEJH12,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==4)$PEJH12,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==5)$PEJH12,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==6)$PEJH12,
        main="Porcentaje Escolaridad de Jefes de hogar con menos de 12",
        xlab="Cluster",
        ylab="Porcentaje")
```

```{r}
boxplot(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==1)$TII,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==2)$TII,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==3)$TII,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==4)$TII,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==5)$TII,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==6)$TII,
        main="Tasa inmigración Intrametropolitana zonal",
        xlab="Cluster",
        ylab="Porcentaje")
```
```{r}
boxplot(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==1)$PHZ,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==2)$PHZ,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==3)$PHZ,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==4)$PHZ,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==5)$PHZ,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==6)$PHZ,
        main="Hacinamiento Zonal",
        xlab="Cluster",
        ylab="Porcentaje")
```
```{r}
boxplot(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==1)$H,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==2)$H,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==3)$H,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==4)$H,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==5)$H,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==6)$H,
        main="Diversidad de usos de suelo",
        xlab="Cluster",
        ylab="H (adimensional)")
```

```{r}
boxplot(subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==1)$DBI,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==2)$DBI,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==3)$DBI,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==4)$DBI,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==5)$DBI,
        subset(select(ANALISIS,PEJH12,TII,PHZ,H,DBI),CLUSTER==6)$DBI,
        main="Distancia a los centros",
        xlab="Cluster",
        ylab="kilometros")
```

